---
title: Project2
output: html_notebook
student name: Lam Phan
Class: Machine Learning
Professor: Karen Mazidi
---

The first data set I took from kaggle
Project: Home Credit Default Risk
Here is the website to get the dataset: https://www.kaggle.com/c/home-credit-default-risk/data



The first data set I take from kaggle
Project: 
here is the link to get the data set:
For this data I would like the machine to learn about the loan type that people are getting base on there income, credit balance, etc. 
```{r}
df1_test <- read.csv(file = "/Users/lamphan/Desktop/MachineLearning Project2/application_test.csv", header=TRUE, sep=",")
df1_train <- read.csv(file = "/Users/lamphan/Desktop/MachineLearning Project2/application_train.csv", header=TRUE, sep=",")
```

```{r}
#check out the features and their data types for the test set
summary(df1_test)
names(df1_test)
str(df1_test)
head(df1_test)
tail(df1_test)
dim(df1_test)
```
```{r}
#check out the features and their data types for the train set
summary(df1_train)
names(df1_train)
str(df1_train)
head(df1_train)
tail(df1_train)
dim(df1_train)
```
```{r}
#data cleaning for test set and train set
test1 <- df1_test[,c("NAME_CONTRACT_TYPE","CODE_GENDER", "FLAG_OWN_CAR", "FLAG_OWN_REALTY", "AMT_INCOME_TOTAL", "AMT_CREDIT","AMT_ANNUITY", "AMT_GOODS_PRICE", "NAME_INCOME_TYPE", "NAME_EDUCATION_TYPE","DAYS_EMPLOYED")]
train1 <- df1_train[,c("NAME_CONTRACT_TYPE","CODE_GENDER", "FLAG_OWN_CAR", "FLAG_OWN_REALTY", "AMT_INCOME_TOTAL", "AMT_CREDIT","AMT_ANNUITY", "AMT_GOODS_PRICE", "NAME_INCOME_TYPE", "NAME_EDUCATION_TYPE","DAYS_EMPLOYED")]
```

```{r}
#check the dataset for train data set again after cleaning
summary(train1)
names(train1)
str(train1)
head(train1)
tail(train1)
dim(train1)
```
```{r}
#check the dataset for test data set again after cleaning
summary(test1)
names(test1)
str(test1)
head(test1)
tail(test1)
dim(test1)
```
For next I would like to use Logistic Regression for this dataset.

```{r}
glm_loan <- glm(NAME_CONTRACT_TYPE~CODE_GENDER +FLAG_OWN_CAR +FLAG_OWN_REALTY + AMT_INCOME_TOTAL + AMT_CREDIT + AMT_ANNUITY + AMT_GOODS_PRICE , data=train1, family=binomial)
summary(glm_loan)
```
```{r}
is.na(glm_loan)
```
```{r}
probs <- predict(glm_loan, newdata=test1, type="response")
probs[is.na(probs)] <- probs[1]
pred_glm_loan <- ifelse(probs>0.5, 2, 1)
```
some NA values appears giving me a hard time getting the accuracy value
so I decided to replace them NA values.
```{r}
glm_loan_acc <- mean(pred_glm_loan==as.integer(test1$NAME_CONTRACT_TYPE))
print(paste("glm loan accuracy = ", glm_loan_acc))
```
accuracy is 96.75% which is not bad.
```{r}
set.seed(2018)
library(tree)
tree_loan1 <- tree(NAME_CONTRACT_TYPE~., data=train1)
plot(tree_loan1)
text(tree_loan1, cex=0.5, pretty=0.5)
```
```{r}
tree_loan1
```
```{r}
summary(tree_loan1)
tree_pred_loan1 <- predict(tree_loan1, newdata=test1, type="class")
table(tree_pred_loan1, test1$NAME_CONTRACT_TYPE)
acc_tree_loan1 <- mean(tree_pred_loan1==test1$NAME_CONTRACT_TYPE)
print(paste("Tree algorithm for loan data set accuracy = ", acc_tree_loan1))
```
The matrix shows us that it caught 149 errors out of 48078 data.
The accuracy for the tree algorithm is 99.228% which is pretty good pretty accurate. 



Applying Naive Bayes
```{r}
library(e1071)
set.seed(2018)
nb_model_loan <- naiveBayes(NAME_CONTRACT_TYPE~., data = train1)
predict_nb_loan <- predict(nb_model_loan, newdata=test1, type="class")
table(predict_nb_loan, test1$NAME_CONTRACT_TYPE)
loan_bayes_acc <- mean(predict_nb_loan==test1$NAME_CONTRACT_TYPE)
print(paste("Naive Bayes algorithm for loan data set accuracy = ", loan_bayes_acc))

print(data.frame(predict_nb_loan, test1$NAME_CONTRACT_TYPE))
```
The accuracy for Naive Bayes is 49% because Naive Bayes Algorithm does not handle large data set very well.


The Algortihms performance and accuracy from best to worst:
Tree Algorithm > Logistic Regression Algorithm > Naive Bayes Algorithm

The second Data set I pick is also from kaggle: YouTube videos viewCount every hour project
Here is the link to the data set: https://www.kaggle.com/nnqkfdjq/statistics-observation-of-random-youtube-video#video_characteristics_upload.csv
In this dataset I would like to pick the view counts as my target for the machine to learn. 
```{r}
df2 <- read.csv(file = "/Users/lamphan/Desktop/MachineLearning Project2/count_observation_upload.csv", header=TRUE, sep=",")
```
```{r}
#check out the features for second data set 
summary(df2)
names(df2)
str(df2)
head(df2)
tail(df2)
dim(df2)
```
cleaning the data, get the features that are needed for the machine to learn
```{r}
df2_fix<- df2[,c("commentCount","dislikeCount", "likeCount", "viewCount", "Time")]
```

```{r}
#check out the features for second data set 
summary(df2_fix)
names(df2_fix)
str(df2_fix)
head(df2_fix)
tail(df2_fix)
dim(df2_fix)
```

check for na values as I learned from the past data set.
```{r}
any(is.na(df2_fix$commentCount))
any(is.na(df2_fix$dislikeCount))
any(is.na(df2_fix$likeCount))
any(is.na(df2_fix$viewCount))
any(is.na(df2_fix$Time))
```
fix na values for the feature:
```{r}
df2_fix$commentCount[is.na(df2_fix$commentCount)] <- mean(df2_fix$commentCount, na.rm=TRUE)
df2_fix$dislikeCount[is.na(df2_fix$dislikeCount)] <- mean(df2_fix$dislikeCount, na.rm=TRUE)
df2_fix$likeCount[is.na(df2_fix$likeCount)] <- mean(df2_fix$likeCount, na.rm=TRUE)
df2_fix$viewCount[is.na(df2_fix$viewCount)] <- mean(df2_fix$viewCount, na.rm=TRUE)
```
check for na values again before I make further process
```{r}
any(is.na(df2_fix$commentCount))
any(is.na(df2_fix$dislikeCount))
any(is.na(df2_fix$likeCount))
any(is.na(df2_fix$viewCount))
any(is.na(df2_fix$Time))
```
na values now seem to be out of the 
I tried to plot but my laptop fail to do so.
Now apply Linear Regression
```{r}
#set test and train sets
set.seed(2018)
i <- sample(1:nrow(df2_fix), nrow(df2_fix)*0.75, replace = FALSE)
train2 <- df2_fix[i,]
test2 <- df2_fix[-i,]
```
generate the linear regression model
```{r}
lm1 <- lm(viewCount~. , data = train2)
```
```{r}
pred_lm <- predict(lm1, newdata= test2)
cor_lm <- cor(pred_lm, test2$viewCount)
print(paste("correlation for linear regression is : ", cor_lm))
mse_lm <- mean((pred_lm - train2$viewCount)^2)
print(paste("mse for linear regression is : ", mse_lm))
```
The correlation between the linear regression model and the target is 0.903 which is good.
though mse is really high
Now apply Knn on second data set:

```{r}
#convert Time to numeric so that we can scale the data set
is.numeric(df2_fix$Time)
df2_fix$Time <- as.numeric(df2_fix$Time)
is.numeric(df2_fix$Time)

```
```{r}
library(caret)
df2_scaled <- data.frame(scale(df2_fix[,c(1:5)])) # scale all the features
```
Check for na values after scaling
```{r}
any(is.na(df2_scaled$commentCount))
any(is.na(df2_scaled$favoriteCount))
any(is.na(df2_scaled$likeCount))
any(is.na(df2_scaled$viewCount))
any(is.na(df2_scaled$Time))
```
devide into train and test set
then generate the fit model
```{r}
train3 <- df2_scaled[i,]
test3 <- df2_scaled[-i,]
fit <- knnreg(train3[,2:4],train3[,1], k= 10 , replace = TRUE)
```
calculate correlation and the mse
```{r}
pred_knn <- predict(fit, test3[,2:4])
cor_knn <- cor(pred_knn, test3$viewCount)
print(paste("correlation for knn regression is : ", cor_knn))
mse_knn <- mean((pred_knn - test3$viewCount)^2)
print(paste("mse for knn regression is : ", mse_knn))
```
The algorith keeps giving the error. It might be because the data set is too large
Conclusion: Linear Regression is the best for second data set
